Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-11-24 text/ms.tex
Edits based on Jon's comments

% TODO highlight that skew would *increase* the probability of higher
% extinction risk

@@

There are two key reasons why we may find little evidence of ecological black
swans. First, they might not exist in higher taxa. Indeed, the majority of
model fitting and risk forecasting assumes that population dynamics are not
heavy tailed \citep[e.g.][]{brook2006a,dennis2006,knape2012}. Alternatively,
heavy-tailed dynamics might exist, but our ability to detect them requires
further development of statistical tools. One such tool is the generalized
extreme value distribution, which has been applied to environmental data
\citep[e.g.][]{katz2005}. This distribution describes the most extreme event
per time interval (e.g.~heaviest rainfall per year), but requires time series
that are sufficiently long to be condensed into time intervals. Another
statistical tool involves fitting a catastrophic mixture distribution in a
state space model to quantify the probability that population events are
extreme, although this is also data intensive \citep{ward2007}. A third tool is
to compare the support for fits of thin- and heavy-tailed distributions
\citep{halley2002}, but this analysis did not quantify the probability of black
swans or allow for population dynamics.

%TODO: this paragraph is written as if geared towards a methods journal. I suggest starting with a statement of general aim rather than method.
Here we develop a framework for identifying ecological black swans in the
process noise of population dynamics, i.e.\ the stochastic jumps in abundance
from one time step to the next. Our framework quantifies both the magnitude and
probability of heavy-tailed dynamics, allows for a range of population dynamic
models, can incorporate observation uncertainty, and can be easily applied to
abundance time series. We apply our model to \NPops\ populations and estimate
the frequency and strength of black-swan events in population dynamics across
taxonomy; verify some causes of black-swan events; and attempt to identify
characteristics of time series or intrinsic life-history characteristics that
are associated with heavy-tailed dynamics.

@@

By allowing the process noise to be drawn from a Student-t distribution we can
estimate the degree to which the process deviations have heavy tails and are
thereof evidence of black-swan events (Fig.~\ref{fig:didactic}a, b). For
example, at $\nu = 3$, the probability of drawing a value more than five
standard deviations below the mean is $0.02$, whereas the probability of
drawing such a value from a normal distribution is tiny ($2.9\cdot10^{-7}$). As
the value of $\nu$ approaches infinity, the distribution approaches the normal
distribution (Fig.~\ref{fig:didactic}a, b).

@@

We chose weakly informative priors to incorporate our understanding of
plausible population dynamics (\citeauthor{gelman2014} \citeyear{gelman2014};
Figs~\ref{fig:didactic}f--h, \ref{fig:priors}). For $\nu$, we chose an
exponential prior with rate parameter of $0.01$ truncated at values above
two---a slightly less informative prior than suggested by
\citet{fernandez1998}. This prior gives only a \basePriorProbHeavy \%
probability that $\nu < 10$ but constrains the sampling sufficiently to avoid
wandering off towards infinity. In any case, for $\nu > 20$ the t distribution
is almost indistinguishable from the normal distribution
(Fig.~\ref{fig:didactic}). We investigated the sensitivity of our results to
weaker and stronger priors (exponential rate parameter $= 0.005, 0.02$;
Fig.~\ref{fig:priors}, Supporting Information). We used simulated data to test
how easily we could detect $\nu$ given different sample sizes and to ensure we
could recover unbiased parameter estimates from the Gompertz model (Supporting
Information).

@@

@@

We investigated possible covariates of heavy-tailed population dynamics
visually and through multilevel modelling. We plotted characteristics of the
time series ($\sigma$, $\lambda$, $b$, and time-series length) along with two
life-history characteristics (body length and maximum lifespan obtained from
\citet{brook2006a}) against our estimated probability of heavy tails, Pr$(\nu <
10)$. We formally investigated these relationships by fitting beta
regression multilevel models \citep{ferrari2004}. Beta regression is useful
for continuous response data that range between zero and one (e.g.\
probabilities); we fit our models with a logit link as is common for logistic
regression \citep{ferrari2004}. We incorporated standard deviations around the
means for covariates that were derived from Gompertz model parameter
estimates. To account for broad patterns of phylogenetic relatedness, we
allowed for hierarchical intercepts at the taxonomic class, order, and species
level (Supporting Information). We fit our model in Stan with weakly
informative priors on the coefficients \citep{gelman2008d} and variance
parameters (\citeauthor{gelman2006c} \citeyear{gelman2006c}
\citeauthor{gelman2014} \citeyear{gelman2014}; Supporting Information).

@@

%TODO: where did this threshold come from?  (nu < 10)
We found strong, but rare, evidence for black-swan population dynamics. By
defining black-swan dynamics as a greater than $0.5$ probability that $\nu <
10$, our main Gompertz model found evidence for heavy tails most frequently
for birds (\birdPH\%) followed by mammals (\mammalsPH\%), and insects
(\insectsPH\%) (Fig.~\ref{fig:nu-coefs}, Table~\ref{tab:causes-supp}). Black
swans were taxonomically widespread, occurring in \POrdersHeavy\% of taxonomic
orders. Accounting for time series length and partially pooling inference
across taxonomic class and order with a multilevel model, there was stronger
evidence for black swans in insect populations than is visually apparent in
Fig.~\ref{fig:nu-coefs}---four of 10 orders with the highest median
probability of heavy tails were insect orders---however, there was
considerable uncertainty in these estimates (Fig.~\ref{fig:posteriors}a).
%TODO: these multilevel modelling results (and fig 3, 4) are out of order

@@

We examined all time series with published explanations of why the black-swan
events occurred (Tables~\ref{tab:sparks} and \ref{tab:causes-supp}). The
majority (\percBSDown \%) of documented events were downward black swans and
%TODO: what else would cause a black swan event? How defined "unfortunate"?
involved a series of unfortunate events. For example, a shortage of nest sites
%TODO how about just calling it a piscivorous waterfowl or something like that
for the shag long-necked bird (\textit{Phalacrocorax aristotelis}) in the
%TODO why did this make the population more vulnerable? Seems like it wouldn't make it more vulnerable I guess I am trying to figure out if these are additive or synergistic. I am guessing you want to avoid those terms, but I might just be more straight with language and avoid the "more vulnerable" phrase.
United Kingdom reduced population productivity and made the population more
vulnerable when a red tide hit in 1968 \citep{potts1980}. The population
experienced a black-swan downswing, but the new availability of nest sites
resulted in a rapid population upswing \citep{potts1980}. As another example,
% TODO true interaction term or just synchronous events?
an interaction between environmental- and predation-mediated population cycles
are thought to have caused a downward black-swan event for a water vole
(\textit{Arvicola terrestris}) population \citep{saucy1994}. Other black swans
were the result of a sequence of extreme climate events on their own. For
instance, severe winters in 1929, 1940--1942, and 1962--1963 were associated
with black-swan downswings in grey heron (\textit{Ardea cinerea}) abundance in
the United Kingdom \citep{stafford1971}. Our analysis finds that the last event
was a combination of two black-swan events in a row and it took the population
three times longer to recover than predicted \citep{stafford1971}.
\textit{(TODO: substitute a parasite example?)}

@@

We found strong evidence for black swans (heavy-tailed process noise) in 3--5\%
% TODO but the students t is a symmetric distribution is that cool?
of ecological time series. Black swans were usually downward events and were
detected more frequently in longer time series and in populations with a
smaller magnitude of process noise. Black swans were not associated with
density dependence, population growth rate, or lifespan. In verified cases,
% TODO 'interaction' word
black-swan events were often a result of the interaction between elements of
extreme climate, predators and parasite cycles, and strong changes in human
pressures. Our empirical results, sensitivity analyses, and simulation tests
suggest that estimating the tail shape of process noise is a viable method of
detecting black-swan population dynamics and if anything will underestimate the
probability of black-swan events. The presence of black swans highlights the
importance of developing management strategies that detect quickly, respond to,
and are robust to extremes in population dynamics---particularly as the
frequency and magnitude of climatic extremes increase over the next century
\citep{easterling2000,ipcc2012}.

@@

We might expect to observe black-swan dynamics in ecological time series
because of unmodelled intrinsic properties of populations or extrinsic forces
acting on populations. Since a t-distribution can be formed by a mixture of
normal distributions \citep[with the same mean and inverse-gamma-distributed
variances,][]{gelman2014}, we could observe heavy tails if we miss some
underlying mixture of intrinsic processes \citep{allen2001}. That process
might be an aggregation of populations across space, or population diversity,
or an intrinsic change in population variability through time. Extrinsic
forces could also cause black-swan dynamics \citep[e.g.][]{nunez2012}. These
forces could be extreme themselves. For example, extreme climate, predation
from (or competition with) other species experiencing black swans, or sharp
changes in human pressure such hunting, fishing, or habitat destruction might
cause black swans. Alternatively, the interaction of multiple ``normal''
extrinsic forces could give rise to black-swan ecological dynamics. This could
occur if the interaction is synergistic \citep[e.g.][]{kirby2009} or even if
non-synergistic interactions experience a rare alignment \citep{denny2009}.

@@

\bibliographystyle{ecologyletters}
\bibliography{/Users/seananderson/Dropbox/tex/jshort,/Users/seananderson/Dropbox/tex/ref3}
%\bibliography{ms}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/t-nu-eg2.pdf}
\caption{
An illustration of fitting population dynamic models that allow for heavy
tails, represented by the Student-t degrees of freedom parameter $\nu$. (a, b)
The probability density for t distributions with a scale parameter of 1 and
different values of $\nu$. Small values of $\nu$ create heavy tails. As $\nu$
approaches infinity the distribution approaches the normal distribution. For
example, at $\nu = 2$, the probability of drawing a value more than five
standard deviations below the mean is 1.8\%, whereas the probability of drawing
such a value from a normal distribution is nearly zero ($2.9\cdot10^{-5}$\%).
(c--e) Simulated population dynamics from a Gompertz model with process noise
drawn from t distributions with three different values of $\nu$. Coloured dots
in panels c and d represent jumps with less than a 1 in 1000 chance of
occurring in a normal distribution. (f--h) Estimates of $\nu$ from models fit
to the times series in panels c--e. Shown are the posterior samples
(histograms), median and interquartile range of the posterior (IQR) (dots and
line segments), and the exponential prior on $\nu$ (dashed lines). Colour
shading behind panels f--h illustrates the region of heavy tails.
}
\label{fig:didactic}
\end{center}
\end{figure}

@@

\caption{Estimates of population dynamics heavy-tailedness for \nuCoefPopN\
  populations of birds, mammals, insects, and fishes. Small values of $\nu$
  ($\lesssim 10$) suggest heavy-tailed black-swan dynamics; larger values of
  $\nu$ suggest approximately normal-tailed dynamics. Vertical points and line
  segments represent posterior medians and 50\% / 90\% credible intervals for
  individual populations. Inset plots show probability that $\nu < 10$
  (probability of heavy tails) for populations arranged by taxonomic order and
  sorted by decreasing mean Pr($\nu < 10$). Taxonomic orders with three or
  fewer populations in panel a are omitted for space. Red to yellow points
  highlight populations with a high to moderately high probability of
  heavy-tailed black-swan dynamics.}

@@

@@

\caption{Potential covariates of heavy-tailed population dynamics (indicated
  by a high probability that $\nu < 10$). Shown are (a--c) parameters from the
  Gompertz heavy-tailed population model ($\sigma$, $\lambda$, $b$), (b)
  number of time steps, (c) body length, and (d) lifespan. For the Gompertz
  parameters, $\sigma$ refers to the scale parameter of the Student-t
  process-noise distribution, $\lambda$ refers to the expected log abundance
  at the next time step at an abundance of one, $b$ refers to the density
  dependence parameter ($1$ is maximally density independent, $0$ is maximally
  density dependent, and $<0$ is inversely density dependent). Circles
  representing a few sharks, crustaceans, and gastropods are filled in white.
  Median and 90\% credible interval posterior predictions of a beta regression
  multilevel model are shown in panels a and d where there was a high
  probability the slope coefficient was different from zero
  (Fig.~\ref{fig:posteriors}b).}

@@

\caption{Posterior probability distributions from beta regression multilevel
  models. (a) Taxonomic-order-level posterior densities of Pr($\nu < 10$)
  (approximately the probability of heavy tails) after accounting for
  time-series length. Estimates are at the geometric mean of time series
  length across all the data (approximately 27 time steps). Colour shading
  refers to taxonomic class (yellow: fishes, green: insects, purple: birds,
  and red: mammals). Dotted vertical line in panel a indicates the median
  expected Pr($\nu < 10$) from the prior distribution. (b) Main effect
  posterior densities for potential covariates of Pr($\nu < 10$). The beta
  regression models were fit on a logit scale with hierarchical intercepts for
  taxonomic class, taxonomic order, and species. All covariates were
  standardized by subtracting their mean and dividing by twice their standard
  deviation. In both panels, short vertical line segments within the density
  polygons indicate median posterior estimates.}

>>>>

@@

There are two key reasons why we may find little evidence of ecological black
swans. First, they might not exist in higher taxa. Indeed, the majority of
model fitting and risk forecasting assumes that population dynamics are normal
tailed on a log scale \citep[e.g.][]{brook2006a,dennis2006,knape2012}.
Alternatively, black-swan dynamics might exist, but our ability to detect
them requires further development of statistical tools. One such tool is the
generalized extreme value distribution, which has been applied to
environmental data \citep[e.g.][]{katz2005}. This distribution describes the
most extreme event per time interval (e.g.~heaviest rainfall per year), but
requires time series that are sufficiently long to be condensed into time
intervals. Another statistical tool involves fitting a catastrophic mixture
distribution in a state space model to quantify the probability that
population events are extreme, although this is also data intensive
\citep{ward2007}. A third tool is to compare the support for fits of thin- and
heavy-tailed distributions \citep{halley2002}, but this analysis did not
quantify the probability of black swans or allow for population dynamics.

Here, we assess the frequency and magnitude of black-swan dynamics across
\NPops\ populations from a wide array of taxonomic groups---mostly birds,
mammals, insects, and fishes. We then identify characteristics of time series
or intrinsic life-history characteristics that are associated with the
detection of black-swan events and attempt to verify known causes. To
accomplish this, we develop a framework for identifying heavy-tailed process
noise in population dynamics, i.e.\ whether the largest stochastic jumps in
abundance from one time step to the next are more extreme than typically seen
with a log-normal distribution. Our framework allows for a range of population
dynamic models, can incorporate observation uncertainty, and can be easily
applied to abundance time series.

@@

By allowing the process noise to be drawn from a Student-t distribution we can
estimate the degree to which the process deviations have heavy tails and are
thereof evidence of black-swan events (Fig.~\ref{fig:didactic}a, b). For
example, at $\nu = 3$, the probability of drawing a value more than five
standard deviations below the mean is $0.02$, whereas the probability of
drawing such a value from a normal distribution is tiny ($2.9\cdot10^{-7}$).
As the value of $\nu$ approaches infinity, the distribution approaches the
normal distribution (Fig.~\ref{fig:didactic}a, b). While, across populations,
the extreme process deviations might tend to be more frequently upwards or
downwards events, a small random sample from a heavy-tailed distribution can
have many outliers on one side and appear asymmetric \citep{gelman2013a}.
Therefore, fitting a symmetric t distribution is appropriate for the
relatively short population time series in our dataset and is agnostic towards
the detection of upwards or downwards black swans.

@@

We chose weakly informative priors to incorporate our understanding of
plausible population dynamics (\citeauthor{gelman2014} \citeyear{gelman2014};
Figs~\ref{fig:didactic}f--h, \ref{fig:priors}). For $\nu$, we chose an
exponential prior with rate parameter of $0.01$ truncated at values above
two---a slightly less informative prior than suggested by
\citet{fernandez1998}. This prior gives only a \basePriorProbHeavy \%
probability that $\nu < 10$ but constrains the sampling sufficiently to avoid
wandering off towards infinity. In any case, for $\nu > 20$ the t distribution
is almost indistinguishable from the normal distribution
(Fig.~\ref{fig:didactic}). Based on the shape of the t distribution, we chose
the probability of $\nu < 10$, Pr($\nu < 10$), to define the probability of
heavy-tailed dynamics. When categorizing a population as heavy or normal tailed,
we used a threshold of 0.5 probability.

@@

In addition to alternative population models, we investigated the sensitivity
of our results to weaker and stronger priors (exponential rate parameter $=
0.005, 0.02$; Fig.~\ref{fig:priors}, Supporting Information). We used
simulated data to test how easily we could detect $\nu$ given different sample
sizes and to ensure we could recover unbiased parameter estimates from the
Gompertz model (Supporting Information).

@@

We investigated possible covariates of heavy-tailed population dynamics
visually and through multilevel modelling. We plotted characteristics of the
time series ($\sigma$, $\lambda$, $b$, and time-series length) along with two
life-history characteristics (body length and maximum lifespan obtained from
\citet{brook2006a}) against our estimated probability of heavy tails, Pr$(\nu <
10)$. We formally investigated these relationships by fitting beta
regression multilevel models \citep{ferrari2004}. The beta probability
distribution can represent continuous response data that range between zero
and one; we fit our models with a logit link as is common for beta and logistic
regression \citep{ferrari2004}. We incorporated standard deviations around the
means for covariates that were derived from Gompertz model parameter
estimates. To account for broad patterns of phylogenetic relatedness, we
allowed for hierarchical intercepts at the taxonomic class, order, and species
level (Supporting Information). We fit our model in Stan with weakly
informative priors on the coefficients \citep{gelman2008d} and variance
parameters (\citeauthor{gelman2006c} \citeyear{gelman2006c}
\citeauthor{gelman2014} \citeyear{gelman2014}; Supporting Information).

@@

We found strong, but rare, evidence for black-swan population dynamics. By
defining black-swan dynamics as a greater than $0.5$ probability that $\nu <
10$, our main Gompertz model found evidence for heavy tails most frequently
for birds (\birdPH\%) followed by mammals (\mammalsPH\%), and insects
(\insectsPH\%) (Fig.~\ref{fig:nu-coefs}, Table~\ref{tab:causes-supp}). Black
swans were taxonomically widespread, occurring in \POrdersHeavy\% of taxonomic
orders. Accounting for time series length and partially pooling inference
across taxonomic class and order with a multilevel model (Supporting
Information), there was stronger evidence for black swans in insect
populations than is visually apparent in Fig.~\ref{fig:nu-coefs}---four of 10
orders with the highest median probability of heavy tails were insect
orders---however, there was considerable uncertainty in these estimates
(Fig.~\ref{fig:posteriors}a).

@@

We examined all time series with published explanations of why the black-swan
events occurred (Tables~\ref{tab:sparks} and \ref{tab:causes-supp}). The
majority of documented events (\percBSDown \%) were downward black swans and
involved a combination of multiple factors. For example, a synchronization of
environmental- and predation-mediated population cycles are thought to have
caused a downward black-swan event for a water vole (\textit{Arvicola
  terrestris}) population \citep{saucy1994}. Other black swans were the result
of a sequence of extreme climate events on their own. For instance, severe
winters in 1929, 1940--1942, and 1962--1963 were associated with black-swan
downswings in grey heron (\textit{Ardea cinerea}) abundance in the United
Kingdom \citep{stafford1971}. Our analysis finds that the last event was a
combination of two black-swan events in a row and it took the population three
times longer to recover than predicted \citep{stafford1971}. Downwards black
swans were sometimes followed by upwards black swans. For example, during a
period of population crowding and nest shortages, a population of European
shag cormorants (\textit{Phalacrocorax aristotelis}) on the Farne Islands, UK,
declined suddenly following a red tide event in 1968 \citep{potts1980}. This
freed quality nest sites for first-time breeders, productivity rapidly
increased, and the population experienced a rapid upswing in abundance.
\citep{potts1980}.

@@

We found strong evidence for black swans (heavy-tailed process noise) in
\overallMinPerc--\overallMaxPerc\% of ecological time series. Black swans were
usually (\percBSDown \%) downward events and were detected more frequently in
longer time series and in populations with a smaller magnitude of process
noise. Black swans were not associated with density dependence, population
growth rate, or lifespan. In verified cases, black-swan events were often a
result of the synchrony of extreme climate, predators and parasite cycles, and
strong changes in human pressures. Our empirical results, sensitivity
analyses, and simulation tests suggest that estimating the tail shape of
process noise is a viable method of detecting black-swan population dynamics
and if anything will underestimate the probability of black-swan events. The
presence of black swans highlights the importance of developing management
strategies that detect quickly, respond to, and are robust to extremes in
population dynamics---particularly as the frequency and magnitude of climatic
extremes increase over the next century \citep{easterling2000,ipcc2012}.

@@

We might expect to observe black-swan dynamics in ecological time series
because of unmodelled intrinsic properties of populations or extrinsic forces
acting on populations. Since a t-distribution can be formed by a mixture of
normal distributions \citep[with the same mean and inverse-gamma-distributed
variances,][]{gelman2014}, we could observe heavy tails if we miss some
underlying mixture of intrinsic processes \citep{allen2001}. That process
might be an aggregation of populations across space, or population diversity,
or an intrinsic change in population variability through time. Extrinsic
forces could also cause black-swan dynamics \citep[e.g.][]{nunez2012}. These
forces could be extreme themselves. For example, extreme climate, predation
from (or competition with) other species experiencing black swans, or sharp
changes in human pressure such hunting, fishing, or habitat destruction might
cause black swans. Alternatively, the synchrony of multiple ``normal''
extrinsic forces could give rise to black-swan ecological dynamics. This could
occur if with a synergistic interaction \citep[e.g.][]{kirby2009} or even if
non-synergistic forces experience a rare alignment \citep{denny2009}.

@@

\bibliographystyle{ecologyletters}
\bibliography{/Users/seananderson/Dropbox/tex/jshort,/Users/seananderson/Dropbox/tex/ref3}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../analysis/t-nu-eg2.pdf}
\caption[An illustration of fitting population dynamic models that allow for heavy
tails, represented by the Student-t degrees of freedom parameter $\nu$.]{
An illustration of fitting population dynamic models that allow for heavy
tails, represented by the Student-t degrees of freedom parameter $\nu$. (a, b)
The probability density for t distributions with a scale parameter of 1 and
different values of $\nu$. Small values of $\nu$ create heavy tails. As $\nu$
approaches infinity the distribution approaches the normal distribution. For
example, at $\nu = 2$, the probability of drawing a value more than five
standard deviations below the mean is 1.8\%, whereas the probability of drawing
such a value from a normal distribution is nearly zero ($2.9\cdot10^{-5}$\%).
(c--e) Simulated population dynamics from a Gompertz model with process noise
drawn from t distributions with three different values of $\nu$. Coloured dots
in panels c and d represent jumps with less than a 1 in 1000 chance of
occurring in a normal distribution. (f--h) Estimates of $\nu$ from models fit
to the times series in panels c--e. Shown are the posterior samples
(histograms), median and interquartile range of the posterior (IQR) (dots and
line segments), and the exponential prior on $\nu$ (dashed lines). Colour
shading behind panels f--h illustrates the region of heavy tails.
}
\label{fig:didactic}
\end{center}
\end{figure}

@@

\caption[Estimates of population dynamics heavy-tailedness for \nuCoefPopN\
  populations of birds, mammals, insects, and fishes.]{Estimates of population dynamics heavy-tailedness for \nuCoefPopN\
  populations of birds, mammals, insects, and fishes. Small values of $\nu$
  ($\lesssim 10$) suggest heavy-tailed black-swan dynamics; larger values of
  $\nu$ suggest approximately normal-tailed dynamics. Vertical points and line
  segments represent posterior medians and 50\% / 90\% credible intervals for
  individual populations. Inset plots show probability that $\nu < 10$
  (probability of heavy tails) for populations arranged by taxonomic order and
  sorted by decreasing mean Pr($\nu < 10$). Taxonomic orders with three or
  fewer populations in panel a are omitted for space. Red to yellow points
  highlight populations with a high to moderately high probability of
  heavy-tailed black-swan dynamics.}

@@

\caption[Posterior probability distributions from beta regression multilevel
  models.]{Posterior probability distributions from beta regression multilevel
  models. (a) Taxonomic-order-level posterior densities of Pr($\nu < 10$)
  (approximately the probability of heavy tails) after accounting for
  time-series length. Estimates are at the geometric mean of time series
  length across all the data (approximately 27 time steps). Colour shading
  refers to taxonomic class (yellow: fishes, green: insects, purple: birds,
  and red: mammals). Dotted vertical line in panel a indicates the median
  expected Pr($\nu < 10$) from the prior distribution. (b) Main effect
  posterior densities for potential covariates of Pr($\nu < 10$). The beta
  regression models were fit on a logit scale with hierarchical intercepts for
  taxonomic class, taxonomic order, and species. All covariates were
  standardized by subtracting their mean and dividing by twice their standard
  deviation. In both panels, short vertical line segments within the density
  polygons indicate median posterior estimates.}

@@

\caption[Potential covariates of heavy-tailed population dynamics (indicated
  by a high probability that $\nu < 10$).]{Potential covariates of heavy-tailed population dynamics (indicated
  by a high probability that $\nu < 10$). Shown are (a--c) parameters from the
  Gompertz heavy-tailed population model ($\sigma$, $\lambda$, $b$), (b)
  number of time steps, (c) body length, and (d) lifespan. For the Gompertz
  parameters, $\sigma$ refers to the scale parameter of the Student-t
  process-noise distribution, $\lambda$ refers to the expected log abundance
  at the next time step at an abundance of one, $b$ refers to the density
  dependence parameter ($1$ is maximally density independent, $0$ is maximally
  density dependent, and $<0$ is inversely density dependent). Circles
  representing a few sharks, crustaceans, and gastropods are filled in white.
  Median and 90\% credible interval posterior predictions of a beta regression
  multilevel model are shown in panels a and d where there was a high
  probability the slope coefficient was different from zero
  (Fig.~\ref{fig:posteriors}b).}

@@

### 2014-11-24 text/sparks.tex
Edits based on Jon's comments

\caption{Example population dynamic black swans from the Global Population
  Dynamics Database and a description of their causes. Red and blue dots
  indicate downward and upward events that have a $10 \cdot 10^{-4}$ probability or less of
  occurring if the population dynamics were explained by a Gompertz model
  with normally distributed process noise. These populations are
  a sample from the heavy-tailed populations we could verify
  (Table~\ref{tab:causes-supp}).}\\

>>>>

\caption[Example population dynamic black swans from the Global Population
  Dynamics Database and a description of their causes.]{Example population dynamic black swans from the Global Population
  Dynamics Database and a description of their causes. Red and blue dots
  indicate downward and upward events that have a $10 \cdot 10^{-4}$ probability or less of
  occurring if the population dynamics were explained by a Gompertz model
  with normally distributed process noise. These populations are
  a sample from the heavy-tailed populations we could verify
  (Table~\ref{tab:causes-supp}).}\\
